SWMPr: An R Package for Retrieving, Organizing, and Analyzing Environmental Data for Estuaries

### 2015-09-02 beck.tex
working version following R journal wrapper file

\documentclass[a4paper]{report}\usepackage[]{graphicx}\usepackage{xcolor}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{RJournal}
\usepackage{amsmath,amssymb,array}
\usepackage{booktabs}

%% load any required packages

% bold/italic text
\newcommand{\Bigtxt}[1]{\textbf{\textit{#1}}}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

%% do not edit, for illustration only
\sectionhead{Contributed research article}
\volume{XX}
\volnumber{YY}
\year{20ZZ}
\month{AAAA}

\begin{article}

%knitr options

\title{SWMPr: An R Package for Retrieving, Organizing, and Analyzing Environmental Data for Estuaries}
\author{by Marcus W Beck}
% ORISE Research Participation Program, USEPA National Health and Environmental Effects Research Laboratory, Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32651, USA

@@

\section*{Abstract}
Standardized monitoring programs have vastly improved the quantity and quality of data that form the basis of environmental decision-making. One example in the United States is the System-Wide Monitoring Program (SWMP) that was implemented in 1995 by the federally-funded National Estuarine Research Reserve System. This program has provided two decades of continuous monitoring data at over 140 fixed stations in 28 estuaries. SWMP data have been used in a variety of applications with the general objective of describing dynamics of estuarine ecosystems to better inform effective coastal management.  However, simple tools for processing and evaluating the increasing quantity of data provided by the monitoring network have prevented broad-scale comparisons between systems and, in some cases, simple trend analysis of water quality parameters at individual sites.  I describe SWMPr, an open-source R package, for use with SWMP environmental data.  The package provides several functions that facilitate data retrieval, organization, and analysis of time series data to describe water quality, weather, and nutrient dynamics in the reserve estuaries.  Previously unavailable functions for estuaries are also provided to estimate rates of ecosystem metabolism using the open-water method.  Tools included with the SWMPr package have facilitated a cross-reserve comparison of trends, including simple evaluation of changes over time and comparisons of patterns in primary productivity.  Overall, the package provides an effective approach to link quantitative information with analysis tools that will greatly inform management programs aimed at coastal protection and restoration.

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> install.packages('SWMPr')
> library(SWMPr)
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # retrieve metadata for all sites
> site_codes()
> 
> # retrieve metadata for a single site
> site_codes_ind('apa')
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # all parameters for a station, most recent
> all_params('hudscwq')
> 
> # get all parameters within a date range
> all_params_dtrng('hudscwq', c('09/10/2012', '02/8/2013'))
> 
> # get single parameter within a date range
> all_params_dtrng('hudscwq', c('09/10/2012', '02/8/2013'), 
+   param = 'do_mgl')
> 
> # single parameter for a station, most recent
> single_param('hudscwq', 'do_mgl')
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # import local data for apaebmet
> 
> # this is an example path with the decompressed csv files
> path <- 'C:/my_path'
> 
> # this is an example path for zipped csv files
> path <- 'C:/my_path.zip'
> 
> # import, do not include file extension
> import_local(path, 'apadbwq') 
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # import all paramaters for the station
> # three most recent records
> exdat <- all_params('apadbwq', Max = 3, trace = F)
> exdat
##         datetimestamp temp f_temp spcond f_spcond sal f_sal do_pct
## 1 2015-08-18 07:00:00   28      0     34        0  21     0     77
## 2 2015-08-18 07:15:00   28      0     34        0  21     0     75
## 3 2015-08-18 07:30:00   28      0     33        0  21     0     82
##   f_do_pct do_mgl f_do_mgl depth f_depth ph f_ph turb f_turb chlfluor
## 1        0      5        0     2       0  8    0   13      0       NA
## 2        0      5        0     2       0  8    0    9      0       NA
## 3        0      6        0     2       0  8    0   11      0       NA
##   f_chlfluor level f_level cdepth clevel f_cdepth f_clevel
## 1         -2    NA      -1     NA     NA       -1         
## 2         -2    NA      -1     NA     NA       -1         
## 3         -2    NA      -1     NA     NA       -1
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # import sample data from package
> data(apadbwq)
> dat <- apadbwq
> 
> # view all attributes of dat
> attributes(dat)
>  
> # view a single attribute of dat
> attr(dat, 'station')
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # view available methods for swmpr class
> methods(class = 'swmpr')
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # qaqc screen for a swmpr object, retain only '0'
> qaqc(dat)
> 
> # retain all data regardless of flag
> qaqc(dat, qaqc_keep = NULL)
> 
> # retain only '0' and '-1' flags
> qaqc(dat, qaqc_keep = c(0, -1))
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # view the number of observations in each QAQC flag
> qaqcchk(dat)
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # get nutrient data, filter qaqc flags
> data(apacpnut)
> dat <- apacpnut
> dat <- qaqc(dat)
> 
> # remove replicate nutrient data
> rem_reps(dat)
> 
> # use different function to aggregate replicates
> func <- function(x) max(x, na.rm = T)
> rem_reps(dat, FUN = func)
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # import data
> data(apaebmet)
> dat <- apaebmet
> 
> # select two parameters from dat
> subset(dat, select = c('rh', 'bp'))
> 
> # subset records greater than or equal to a date
> subset(dat, subset = '2013-01-01 0:00', operator = '>=')
> 
> # subset records within a date range
> subset(dat, subset = c('2012-07-01 6:00', '2012-08-01 18:15'))
> 
> # subset records within a date range, select two parameters
> subset(dat, subset = c('2012-07-01 6:00', '2012-08-01 18:15'),
+   select = c('atemp', 'totsorad'))
> 
> # remove rows/columns that do not contain data
> subset(dat, rem_rows = T, rem_cols = T)
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # import, qaqc removal
> data(apadbwq)
> dat <- qaqc(apadbwq)
> 
> # convert time series to two hour invervals
> # tolerance of +/- 30 minutes for matching existing data
> setstep(dat, timestep = 120, differ = 30)
> 
> # convert a nutrient time series to a continuous time series
> # then remove empty rows and columns
> data(apacpnut)
> dat_nut <- apacpnut
> dat_nut <- setstep(dat_nut, timestep = 60)
> subset(dat_nut, rem_rows = T, rem_cols = T)
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # get nut, wq, and met data as separate objects
> data(apacpnut)
> data(apacpwq)
> data(apaebmet)
> swmp1 <- apacpnut
> swmp2 <- apacpwq
> swmp3 <- apaebmet
> 
> # combine nut and wq data by union
> comb(swmp1, swmp2, method = 'union')
> 
> # combine nut and wq data by intersect
> comb(swmp1, swmp3, method = 'intersect')
> 
> # combine nut, wq, and met data by nut time series, two hour time step
> comb(swmp1, swmp2, swmp3, timestep = 120, method = 'apacpnut')
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # get data, keep all observations
> data(apacpnut)
> dat <- qaqc(apacpnut, qaqc_keep = NULL)
> 
> # aggregate by quarters
> agg_dat <- aggreswmp(dat, by = 'quarters')
> nrow(agg_dat)
## [1] 47
> # aggregate by quarters, remove rows with NA values
> # note the reduction in the number of rows
> agg_dat2 <- aggreswmp(dat, by = 'quarters', na.action = na.omit)
> nrow(agg_dat2)
## [1] 16
\end{verbatim}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # import data, qaqc and subset
> data(apadbwq)
> dat <- qaqc(apadbwq)
> dat <- subset(dat, select = 'do_mgl', 
+   subset = c('2012-07-09 00:00', '2012-07-24 00:00')
+   )
> 
> # smooth using a window of 50 observations
> dat_smooth <- smoother(dat, window = 50, params = 'do_mgl')
> 
> # plot raw and smoothed
> plot(dat)
> lines(dat_smooth, col = 'red', lwd = 2)
\end{verbatim}
\end{kframe}\begin{figure}[!h]

@@

\end{knitrout}

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # get data, qaqc and subset
> data(apadbwq)
> dat <- qaqc(apadbwq)
> dat <- subset(dat, select = 'do_mgl', 
+   subset = c('2013-01-22 00:00', '2013-01-26 00:00'))
> 
> # interpolate, maxgap of 10 records
> fill1 <- na.approx(dat, params = 'do_mgl', maxgap = 10)
> 
> # interpolate maxgap of 30 records
> fill2 <- na.approx(dat, params = 'do_mgl', maxgap = 30)
> 
> # plot for comparison
> par(mfrow = c(3, 1))
> plot(dat, main = 'Raw')
> plot(fill1, col = 'red', main = 'Interpolation - maximum gap of 10 records')
> lines(dat)
> plot(fill2, col = 'red', main = 'Interpolation - maximum gap of 30 records')
> lines(dat)
\end{verbatim}
\end{kframe}\begin{figure}[!h]

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # get data
> data(apadbwq)
> dat <- apadbwq
> 
> # subset for daily decomposition
> dat <- subset(dat, subset = c('2013-07-01 00:00', '2013-07-31 00:00'))
> 
> # daily decomposition of DO and plot
> dc_dat <- decomp(dat, param = 'do_mgl', frequency = 'daily')
> plot(dc_dat)
\end{verbatim}
\end{kframe}\begin{figure}[!h]

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # get data
> data(apacpnut)
> dat <- apacpnut
> dat <- qaqc(dat, qaqc_keep = NULL)
> 
> # decomposition of chl
> decomp_cj(dat, param = 'chla_n')
\end{verbatim}
\end{kframe}\begin{figure}[!h]

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> ## import data
> data(apacpnut)
> dat <- qaqc(apacpnut)
> 
> ## plot
> plot_summary(dat, param = 'chla_n', years = c(2007, 2013))
\end{verbatim}
\end{kframe}\begin{figure}[!h]

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> ## import water quality and weather data
> data(apadbwq)
> data(apaebmet)
> 
> ## qaqc, combine
> wq <- qaqc(apadbwq)
> met <- qaqc(apaebmet)
> dat <- comb(wq, met)
> 
> ## estimate metabolism
> res <- ecometab(dat, trace = FALSE)
> plot_metab(res)
\end{verbatim}
\end{kframe}\begin{figure}[!h]

@@

\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{kframe}
\begin{verbatim}
> # plot the stations at Jacques Cousteau reserve
> map_reserve('jac')
\end{verbatim}
\end{kframe}\begin{figure}[!h]

@@

% metabolism figure
\begin{knitrout}\small
\definecolor{shadecolor}{rgb}{1, 1, 1}\color{fgcolor}\begin{figure}[!h]

{\centering \includegraphics[width=0.00\textwidth]{figure/metab_plo-1}

@@

\end{article}

\end{document}

>>>>

\title{SWMPr: An R Package for Retrieving, Organizing, and Analyzing Environmental Data for Estuaries}
\author{by Marcus W Beck}

@@

\abstract{
Standardized monitoring programs have vastly improved the quantity and quality of data that form the basis of environmental decision-making. One example in the United States is the System-Wide Monitoring Program (SWMP) that was implemented in 1995 by the federally-funded National Estuarine Research Reserve System. This program has provided two decades of continuous monitoring data at over 140 fixed stations in 28 estuaries. SWMP data have been used in a variety of applications with the general objective of describing dynamics of estuarine ecosystems to better inform effective coastal management.  However, simple tools for processing and evaluating the increasing quantity of data provided by the monitoring network have prevented broad-scale comparisons between systems and, in some cases, simple trend analysis of water quality parameters at individual sites.  I describe SWMPr, an open-source R package, for use with SWMP environmental data.  The package provides several functions that facilitate data retrieval, organization, and analysis of time series data to describe water quality, weather, and nutrient dynamics in the reserve estuaries.  Previously unavailable functions for estuaries are also provided to estimate rates of ecosystem metabolism using the open-water method.  Tools included with the SWMPr package have facilitated a cross-reserve comparison of trends, including simple evaluation of changes over time and comparisons of patterns in primary productivity.  Overall, the package provides an effective approach to link quantitative information with analysis tools that will greatly inform management programs aimed at coastal protection and restoration.
}

@@

\begin{example}
> install.packages('SWMPr')
> library(SWMPr)
\end{example}

@@

\begin{example}
> # retrieve metadata for all sites
> site_codes()
> 
> # retrieve metadata for a single site
> site_codes_ind('apa')
\end{example}

@@

\begin{example}
> # all parameters for a station, most recent
> all_params('hudscwq')
> 
> # get all parameters within a date range
> all_params_dtrng('hudscwq', c('09/10/2012', '02/8/2013'))
> 
> # get single parameter within a date range
> all_params_dtrng('hudscwq', c('09/10/2012', '02/8/2013'), 
+   param = 'do_mgl')
> 
> # single parameter for a station, most recent
> single_param('hudscwq', 'do_mgl')
\end{example}

@@

\begin{example}
> # import local data for apaebmet
> 
> # this is an example path with the decompressed csv files
> path <- 'C:/my_path'
> 
> # this is an example path for zipped csv files
> path <- 'C:/my_path.zip'
> 
> # import, do not include file extension
> import_local(path, 'apadbwq') 
\end{example}

@@

\begin{example}
> # import all paramaters for the station
> # three most recent records
> exdat <- all_params('apadbwq', Max = 3, trace = F)
> exdat
##         datetimestamp temp f_temp spcond f_spcond sal f_sal do_pct
## 1 2015-08-18 07:00:00   28      0     34        0  21     0     77
## 2 2015-08-18 07:15:00   28      0     34        0  21     0     75
## 3 2015-08-18 07:30:00   28      0     33        0  21     0     82
##   f_do_pct do_mgl f_do_mgl depth f_depth ph f_ph turb f_turb chlfluor
## 1        0      5        0     2       0  8    0   13      0       NA
## 2        0      5        0     2       0  8    0    9      0       NA
## 3        0      6        0     2       0  8    0   11      0       NA
##   f_chlfluor level f_level cdepth clevel f_cdepth f_clevel
## 1         -2    NA      -1     NA     NA       -1         
## 2         -2    NA      -1     NA     NA       -1         
## 3         -2    NA      -1     NA     NA       -1
\end{example}

@@

\begin{example}
> # import sample data from package
> data(apadbwq)
> dat <- apadbwq
> 
> # view all attributes of dat
> attributes(dat)
>  
> # view a single attribute of dat
> attr(dat, 'station')
\end{example}

@@

\begin{example}
> # view available methods for swmpr class
> methods(class = 'swmpr')
\end{example}

@@

\begin{example}
> # qaqc screen for a swmpr object, retain only '0'
> qaqc(dat)
> 
> # retain all data regardless of flag
> qaqc(dat, qaqc_keep = NULL)
> 
> # retain only '0' and '-1' flags
> qaqc(dat, qaqc_keep = c(0, -1))
\end{example}

@@

\begin{example}
> # view the number of observations in each QAQC flag
> qaqcchk(dat)
\end{example}

@@

\begin{example}
> # get nutrient data, filter qaqc flags
> data(apacpnut)
> dat <- apacpnut
> dat <- qaqc(dat)
> 
> # remove replicate nutrient data
> rem_reps(dat)
> 
> # use different function to aggregate replicates
> func <- function(x) max(x, na.rm = T)
> rem_reps(dat, FUN = func)
\end{example}

@@

\begin{example}
> # import data
> data(apaebmet)
> dat <- apaebmet
> 
> # select two parameters from dat
> subset(dat, select = c('rh', 'bp'))
> 
> # subset records greater than or equal to a date
> subset(dat, subset = '2013-01-01 0:00', operator = '>=')
> 
> # subset records within a date range
> subset(dat, subset = c('2012-07-01 6:00', '2012-08-01 18:15'))
> 
> # subset records within a date range, select two parameters
> subset(dat, subset = c('2012-07-01 6:00', '2012-08-01 18:15'),
+   select = c('atemp', 'totsorad'))
> 
> # remove rows/columns that do not contain data
> subset(dat, rem_rows = T, rem_cols = T)
\end{example}

@@

\begin{example}
> # import, qaqc removal
> data(apadbwq)
> dat <- qaqc(apadbwq)
> 
> # convert time series to two hour invervals
> # tolerance of +/- 30 minutes for matching existing data
> setstep(dat, timestep = 120, differ = 30)
> 
> # convert a nutrient time series to a continuous time series
> # then remove empty rows and columns
> data(apacpnut)
> dat_nut <- apacpnut
> dat_nut <- setstep(dat_nut, timestep = 60)
> subset(dat_nut, rem_rows = T, rem_cols = T)
\end{example}

@@

\begin{example}
> # get nut, wq, and met data as separate objects
> data(apacpnut)
> data(apacpwq)
> data(apaebmet)
> swmp1 <- apacpnut
> swmp2 <- apacpwq
> swmp3 <- apaebmet
> 
> # combine nut and wq data by union
> comb(swmp1, swmp2, method = 'union')
> 
> # combine nut and wq data by intersect
> comb(swmp1, swmp3, method = 'intersect')
> 
> # combine nut, wq, and met data by nut time series, two hour time step
> comb(swmp1, swmp2, swmp3, timestep = 120, method = 'apacpnut')
\end{example}

@@

\begin{example}
> # get data, keep all observations
> data(apacpnut)
> dat <- qaqc(apacpnut, qaqc_keep = NULL)
> 
> # aggregate by quarters
> agg_dat <- aggreswmp(dat, by = 'quarters')
> nrow(agg_dat)
## [1] 47
> # aggregate by quarters, remove rows with NA values
> # note the reduction in the number of rows
> agg_dat2 <- aggreswmp(dat, by = 'quarters', na.action = na.omit)
> nrow(agg_dat2)
## [1] 16
\end{example}

@@

\begin{example}
> # import data, qaqc and subset
> data(apadbwq)
> dat <- qaqc(apadbwq)
> dat <- subset(dat, select = 'do_mgl', 
+   subset = c('2012-07-09 00:00', '2012-07-24 00:00')
+   )
> 
> # smooth using a window of 50 observations
> dat_smooth <- smoother(dat, window = 50, params = 'do_mgl')
> 
> # plot raw and smoothed
> plot(dat)
> lines(dat_smooth, col = 'red', lwd = 2)
\end{example}

\begin{figure}[!h]

@@

@@

\begin{example}
> # get data, qaqc and subset
> data(apadbwq)
> dat <- qaqc(apadbwq)
> dat <- subset(dat, select = 'do_mgl', 
+   subset = c('2013-01-22 00:00', '2013-01-26 00:00'))
> 
> # interpolate, maxgap of 10 records
> fill1 <- na.approx(dat, params = 'do_mgl', maxgap = 10)
> 
> # interpolate maxgap of 30 records
> fill2 <- na.approx(dat, params = 'do_mgl', maxgap = 30)
> 
> # plot for comparison
> par(mfrow = c(3, 1))
> plot(dat, main = 'Raw')
> plot(fill1, col = 'red', main = 'Interpolation - maximum gap of 10 records')
> lines(dat)
> plot(fill2, col = 'red', main = 'Interpolation - maximum gap of 30 records')
> lines(dat)
\end{example}

\begin{figure}[!h]

@@

\begin{example}
> # get data
> data(apadbwq)
> dat <- apadbwq
> 
> # subset for daily decomposition
> dat <- subset(dat, subset = c('2013-07-01 00:00', '2013-07-31 00:00'))
> 
> # daily decomposition of DO and plot
> dc_dat <- decomp(dat, param = 'do_mgl', frequency = 'daily')
> plot(dc_dat)
\end{example}

\begin{figure}[!h]

@@

\begin{example}
> # get data
> data(apacpnut)
> dat <- apacpnut
> dat <- qaqc(dat, qaqc_keep = NULL)
> 
> # decomposition of chl
> decomp_cj(dat, param = 'chla_n')
\end{example}

\begin{figure}[!h]

@@

\begin{example}
> ## import data
> data(apacpnut)
> dat <- qaqc(apacpnut)
> 
> ## plot
> plot_summary(dat, param = 'chla_n', years = c(2007, 2013))
\end{example}

\begin{figure}[!h]

@@

\begin{example}
> ## import water quality and weather data
> data(apadbwq)
> data(apaebmet)
> 
> ## qaqc, combine
> wq <- qaqc(apadbwq)
> met <- qaqc(apaebmet)
> dat <- comb(wq, met)
> 
> ## estimate metabolism
> res <- ecometab(dat, trace = FALSE)
> plot_metab(res)
\end{example}

\begin{figure}[!h]

@@

\begin{example}
> # plot the stations at Jacques Cousteau reserve
> map_reserve('jac')
\end{example}

\begin{figure}[!h]

@@

% metabolism figure

\begin{figure}[!h]
{\centering \includegraphics[width=0.00\textwidth]{figure/metab_plo-1}

@@

\address{Marcus W Beck\\
  ORISE Research Participation Program\\
  USEPA National Health and Environmental Effects Research Laboratory, Gulf Ecology Division\\
  1 Sabine Island Drive, Gulf Breeze, FL 32651\\
  USA\\}
\email{beck.marcus@epa.gov}
